---
title: 'Assignment 3: Solution'
author: "Karsten Donnay"
date: "16.6.2021"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    fig_height: 6
    highlight: tango
    theme: spacelab
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
editor_options:
  chunk_output_type: console
---



```{r knitr_init, echo=TRUE, cache=FALSE, warning=FALSE}
library(knitr)

### Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
rm(list = ls())
```

# Preparation

Install the 'nycflights13' package and load the data into R. These data cover the airline on-time data for all flights departing NYC in 2013. It also includes useful 'metadata' on airlines, airports, weather, and planes.
```{r,echo=TRUE}
library(nycflights13)
```

# Overview
You can get a basic overview of the dataset with these functions
```{r,echo=TRUE}
# How many rows and columns?
dim(flights)  #  or: nrow(flights)    ncol(flights)
# What are the names of the variables/columns?
colnames(flights)
# Summary statistics
summary(flights)
```


# Assignment 1: Subsetting and alterations with dplyr

### (a) Create a new variable

Use dplyr to create a variable 'caught_up' that only consists of values that are TRUE or FALSE and which indicates whether a flight *caught up* with a departure delay, i.e. it should be TRUE if the delay at arrival was less than the delay at departure and FALSE otherwise.
 
 
```{r,echo=TRUE}
flights <- dplyr::mutate(flights, caught_up = arr_delay < dep_delay)
```
 
 
### (b) Extraction of observations

Use dplyr to filter the dataset to include only flights that had a delayed departure. Report which percentage of all the flights had a delayed departure. How many of those delayed flights also had a delayed arrival?
 
```{r,echo=TRUE}
del_dep <- dplyr::filter(flights, dep_delay > 0)
print(paste("Overall, ",round(nrow(del_dep)*100/nrow(flights), 1), "% of the flights had a delayed departure."))
del_dep_del_arr <- dplyr::filter(del_dep, arr_delay > 0)
print(paste("Overall, ", round(nrow(del_dep_del_arr)*100/nrow(del_dep), 1), "% of the flights with delayed departure also had a delayed arrival."))
```

 
# Assignment 2: Summary statistics

### (a) Summary statistics 1

Do flights from JFK have a greater departure delay than flights from EWR on average? Use dplyr to find out.

```{r,echo=TRUE}
library(dplyr)
flights %>% dplyr::group_by(origin) %>% dplyr::summarise(avg_dep_delay = mean(dep_delay, na.rm=T))
print("EWR has a higher departure delay (15.1) than JFK (12.1)")
```

### (b) Summary statistics 2
Which NYC airport is the most common for flying to Chicago O'Hare International Airport (ORD)? Use dplyr to find out.


```{r,echo=TRUE}
library(dplyr)
flights %>% dplyr::filter(dest == "ORD") %>% dplyr::group_by(origin) %>% dplyr::summarise(freq = n())
print("LGA is the airport where the most flights to ORD depart from.")
```

# Assignment 3: Rewriting

### Piping

Rewrite the following statement with a pipe operator (%>%).

```{r,echo=TRUE}
library(dplyr)
set.seed(12345)
sum(select(sample_n(filter(flights, origin=="JFK",dest=="PHX"), 200), air_time), na.rm=T)
```

```{r,echo=TRUE}
set.seed(12345)
flights %>% filter(origin == "JFK", dest == "PHX") %>% sample_n(200) %>% select(air_time) %>% sum(na.rm=T)
```

### dplyr and data.table

Write the following statement with dplyr and in data.table format.

 * "Average departure delay for every flight to Phoenix (PHX) differentiated by carrier and airport of origin."

```{r,echo=TRUE}
library(dplyr)
library(data.table)

solution_dplyr <- flights %>% 
  filter(dest == "PHX") %>% 
  group_by(carrier, origin) %>% 
  summarise(avg_dep_delay = mean(dep_delay, na.rm=T)) %>% 
  select(carrier, origin, avg_dep_delay)

solution_dtable <- as.data.table(flights)[dest == "PHX",
                                     .(avg_dep_delay = mean(dep_delay, na.rm=TRUE)),
                                     by = list(carrier, origin)]
```


